---
title: "Practical Machine Learning—Project Course"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries 



```{r comment = "", warning=FALSE, message=FALSE}

library(rpart.plot)
library(caret)

set.seed(12345)
```

# Data Importing

```{r pressure, comment = ""}

train_d <- read.csv('pml-training.csv', header = TRUE)
val_test <- read.csv('pml-testing.csv', header = TRUE)
  

```

## Data Partitioning  

```{r}



inTrain <- createDataPartition(y = train_d$classe, 
                               p = 0.6, 
                               list = FALSE)

training <- train_d[inTrain, ]; testing <- train_d[-inTrain, ]
dim(training); dim(testing)

```

# Data Cleaning 

## Identify and remove near zero variable from training dataset

``` {r}

nzv_mat <- nearZeroVar(training, saveMetrics = TRUE)

var_names <- rownames(nzv_mat[nzv_mat$nzv == TRUE,])

nzv_var <- names(training) %in% var_names


training <- training[!nzv_var]

dim(training)

# removing the first columns as it is not relevant to the analysis
training <- training[c(-1)]
```

## Omitting variables containing  more than 70% NAs

```{r }

newTraining <- training #creating another subset to iterate in loop

for(i in 1:length(training)) { 
  if( sum( is.na( training[, i] ) ) /nrow(training) >= .7 ) { #if n?? NAs > 60% of total observations
    for(j in 1:length(newTraining)) {
      if( length( grep(names(training[i]), names(newTraining)[j]) ) ==1)  { #if the columns are the same:
        newTraining <- newTraining[ , -j] #Remove that column
      }   
    } 
  }
}




training <- newTraining
```

We need to apply the same transformation to testing and validating sets

```{r}

cln_names <- colnames(training)

testing <- testing[cln_names]

```


# Fittinga a classification model

```{r}

# Fitting models 

clas.mod <- rpart(classe ~ ., data = training, method = "class")
# fancyRpartPlot(clas.mod)

```

## Apply the model to the testing dataset and produce a confusion matrix

```{r}
clas.pred<- predict(clas.mod, testing, type = "class")

confusionMatrix(clas.pred, as.factor(testing$classe))

```

## Predict 20 different cases from the validation set
```{r}

pred <- predict(clas.mod, val_test[cln_names[-58]], type = 'class')
print(pred)
```